---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r loading}
pacman:: p_load(rethinking, corrplot)
library(rethinking, bayesplot)

setwd('~/upgraded-lamp')
d <- read.csv('Assignment3Data.csv') 

#only Visit 1
d1 <- subset(d, d$ASD == 1 & d$Visit == 1 & d$ADOS > 0) #subsetting ASD-participants and visit 1 #what is this one kid with zero ADOS doing in our ASD data #go away plz
d2 <- subset(d, d$ASD == 0 & d$Visit == 1 ) #subsetting TD-participants and visit 1 

#For later use
d3 <- subset(d, d$Visit == 1 )


#testing difference in ADOS between the two groups
t.test(d1$ADOS, d2$ADOS)

```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.


The questions you will have to answer are:

01. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report 1) the model, 2) a plot of the model, and 3) a couple of lines describing the quality of the model and interpreting the results.

P.S. Should you scale?
P.P.S. Don't forget to motivate your priors. See: predictions

#1.1. Verbal IQ and ADOS
```{r inspecting data}

dens(d1$VerbalIQ)

```
The data has two peaks, so we might have to perform a transformation.
Log-transforming is an option, when log-transforming, the data suddenly has three peaks, so we decided on just scaling

```{r 1.1. verbal IQ and ADOS: Setting prior}
###SCALING
VerbalIQ_scale <- d1 
VerbalIQ_scale$VerbalIQ <- scale(d1$VerbalIQ) #scaling: it centers the mean around 0 and makes its sd 1
VerbalIQ_scale$ADOS <- scale(d1$ADOS)

###DENSITY PLOT
dens(VerbalIQ_scale$ADOS)

###BUILDING THE MODEL
verbal_m <- rethinking::map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b* VerbalIQ,
    a ~ dnorm (0, 5), #the mean of Verbal IQ is centered around 0 after scaling that is why we set it to (0, 5), the scaling sets the standard deviation to 1 so setting it to 5 makes it incredibly weak, and we don't have (except for the data we are using in this model) any prior knowledge we can use for informing this prior
    b ~ dnorm (-0.1, 5), #what is the change in expected verbal iq when x1 changes by 1 unit -> ß, slope, we expect beta to be negative, because we expect the following the higher the ADOS the lower the Verbal IQ, we set this to a very small number though, because be can't base this knowledge on actual data, but rather our "knowledge" about the world (autism)
    sigma ~ dunif (0, 5) #bounding it to zero, and again a very big standard deviation -> another weak prior
  ), 
  data = VerbalIQ_scale)

#Estimates
precis(verbal_m)


###PLOTTING POSTERIOR WITH ACTUAL DATA
##PLOT 1
plot(ADOS ~ VerbalIQ, data = VerbalIQ_scale)
abline(a=coef(verbal_m)["a"], b=coef(verbal_m)["b"])
##PLOT 2
post <- extract.samples(verbal_m)
post[1:5,] #expecting five rows
#function
N <- 30
dN <- VerbalIQ_scale[ 1:N , ]
# extract 20 samples from the posterior #the MAP line is just the posterior mean, the most plausible line in the universe of infinite lines that the posterior distribution has considered, so plotting uncertainty makes a lot of sense, it could be that there are many lines with nearly the same posterior probability
post <- extract.samples( verbal_m , n=50 )
# display raw data and sample size
plot( dN$VerbalIQ , dN$ADOS,
    xlim=range(VerbalIQ_scale$VerbalIQ) , ylim=range(VerbalIQ_scale$ADOS) ,
    col=rangi2 , xlab="ADOS" , ylab="VerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )
#there is some uncertainity, but nevertheless the realtionship between ADOS and VerbalIQ seems to be always a negative one, so we think our model is good.

modelVerb <- brms::bf("ADOS ~ VerbalIQ")
brm_modverb <- brm(
  modelVerb, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = VerbalIQ_scale, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_modverb, nsamples = 100)
#looking pretty good, the simulated data are somewhat consistent with the actual data, at least approximately




```


#1.2. Non Verbal IQ and ADOS
```{r 1.2. inspecting data}

dens(d1$NonVerbalIQ)

```
This is Gaussian "enough" for us.
Mean around 27 with sd 15, we will scale it though, to be able to set somewhat sensible priors without looking at the data

```{r 1.2. building and plotting model + quality check}

###SCALING 
NonVerbalIQ_scale <- d1
NonVerbalIQ_scale$NonVerbalIQ <- scale(d1$NonVerbalIQ)
NonVerbalIQ_scale$ADOS <- scale(d1$ADOS)
#Densitiy Plot
dens(NonVerbalIQ_scale$ADOS)

###BUILDING THE MODEL
#NonVerbalIQ
nonverbal_m <- rethinking::map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b* NonVerbalIQ,
    a ~ dnorm (0, 5), #same reasons as in model one, see above
    b ~ dnorm (-0.1, 5),
    sigma ~ dunif (0, 5)
  ), 
  data = NonVerbalIQ_scale)

#MAP Estimates
precis(nonverbal_m)

#PLOTTING 
#Plot 1
plot(ADOS ~ NonVerbalIQ, data = NonVerbalIQ_scale)
abline(a=coef(nonverbal_m)["a"], b=coef(nonverbal_m)["b"])

#Plot 2
post <- extract.samples(nonverbal_m)
post[1:5,] #expecting five rows
#function
N <- 30
dN <- NonVerbalIQ_scale[ 1:N , ]
# extract 20 samples from the posterior
post <- extract.samples( nonverbal_m , n=50 )
# display raw data and sample size
plot( dN$NonVerbalIQ , dN$ADOS,
    xlim=range(NonVerbalIQ_scale$NonVerbalIQ) , ylim=range(NonVerbalIQ_scale$ADOS) ,
    col=rangi2 , xlab="NonVerbalIQ" , ylab="ADOS" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )

modelNonVerb <- brms::bf("ADOS ~ NonVerbalIQ")
brm_mod2 <- brm(
  modelNonVerb, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = NonVerbalIQ_scale, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_mod2, nsamples = 100)
#looking good enough for us, estimated data goes a little wild when it comes to the peak but, it looks somewhat like the actual data without being essentially the same, which we wouldn't want


```


```{r SOCIAL IQ & ADOS}

#SCALING
SocialIQ_scale <- d1
SocialIQ_scale$SocialIQ <- scale(d1$SocialIQ)
SocialIQ_scale$ADOS <- scale(d1$ADOS)

#DENSITY PLOT
dens(SocialIQ_scale$ADOS)


#BUILDING THE MODEL
#Social IQ
social_m <- rethinking::map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b* SocialIQ,
    a ~ dnorm (0, 5), #see VerbalIQ model for reasoning about priors
    b ~ dnorm (-0.1, 5), 
    sigma ~ dunif (0, 5)
  ), 
  data = SocialIQ_scale)

#Estimates
precis(social_m)

###PLOTTING
##PLot 1
plot(ADOS ~ SocialIQ, data = SocialIQ_scale)
abline(a=coef(social_m)["a"], b=coef(social_m)["b"])

##Plot2
post <- extract.samples(social_m)
post[1:5,] #expecting five rows
#function
N <- 30
dN <- SocialIQ_scale[ 1:N , ]
#extract 20 samples from the posterior
post <- extract.samples( social_m , n=50 )
#display raw data and sample size
plot( dN$SocialIQ , dN$ADOS,
    xlim=range(SocialIQ_scale$SocialIQ) , ylim=range(SocialIQ_scale$ADOS) ,
    col=rangi2 , xlab="SocialIQ" , ylab="ADOS" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )


#brms model so we can pp_check
modelSocial <- brms::bf("ADOS ~ SocialIQ")
brm_mod3 <- brm(
  modelSocial, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = SocialIQ_scale, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_mod3, nsamples = 100)
#looking good enough, some wildness and different peaks, but okay nok

```


##Task 2
2. Do the different aspects of IQ account for different portions of the variance in ADOS?

2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?

2.2. Build the model, assess its quality, write a few lines interpreting the results.
```{r 2.2. model building}
library(rethinking)
library(brms)
library(bayesplot)

##SCALING #all the variables before building the model
d6 <- d1
d6$VerbalIQ <- base::scale(d1$VerbalIQ)
d6$NonVerbalIQ <- base::scale(d1$NonVerbalIQ)
d6$SocialVerbalIQ <- base::scale(d1$SocialIQ)
d6$ADOS <- scale(d1$ADOS)
d6 <- base::subset(d6, d6$ChildID > 1) #getting rid of that annyoing kid with NAs

##Building the Model with all IQ scores
m4 <- rethinking::map( 
  alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- a + v*VerbalIQ  + nv* NonVerbalIQ + s*SocialIQ , 
a ~ dnorm( 1 , 5 ) ,
v ~ dnorm( -0.1 , 5 ) ,
nv ~ dnorm( -0.1 , 5 ) ,
s ~ dnorm( -0.1, 5 ) ,
sigma ~ dunif( 0 , 5 )
),
data=d6 )

#Model Estimates
precis(m4)
#89% of n's probability lies between -2.86 and -0.55 suggesting that values close to zero and or above are not compatible, it looks like
#s seams to have little effect extremly close to zero (approaching from both side 89% of the probability lie between -0.07 and 0.08) and surprisingly a positive effect ?
#v and nv have negative effects of about the same size

#Visualizing MAP values
plot(precis(m4))


##Model in brms
#Building the model in brm as well, so we can do a beautiful pp_check
model <- brms::bf("ADOS ~ VerbalIQ + NonVerbalIQ + SocialVerbalIQ")
brm_mod8 <- brm(
  model, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = d6, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_mod8, nsamples = 100)
#


```


```{r 2.2 correlation between measures}

##Verbal and Social IQ
cor.test(d1$VerbalIQ, d1$SocialIQ, method = "pearson") #0.5450814 
(0.5450814)^2 #R^2 = 0.2971137

##Verbal and Nonverbal IQ
cor.test(d1$NonVerbalIQ, d1$VerbalIQ, method = "pearson") #0.7085501
(0.7085501)^2 #R^2 = 0.5020432

##Nonverbal and Social IQ
cor.test(d1$NonVerbalIQ, d1$SocialIQ, method = "pearson") #0.396656
(0.396656)^2 #R^2 = 0.157336

#Scatterplot Matrix
rethinking::pairs(~ SocialVerbalIQ + VerbalIQ + NonVerbalIQ , data= d6, 
   main="Simple Scatterplot Matrix")

```
All measure are correlated and especially Verbal IQ and NonVerbal IQ seem to share at least 50% of their variance, explaining why our model looks very similar to the ones in Task 1, where we only used one predictor


#Task 3
3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.
# Yes: ASD is not binary. Some children might lean towards ASD, but be categorized TD.
#ADOS is, afterall, symptom severity. A continunous scale. We are interested in seeing ADOS predicting IQ [thus, IQ ~ ADOS]

We certainly expect a relationship between IQ and ADOS, however we might run into problems because the ADOS scale is limited to >= 0, so almost every TD kid will end up at ADOS = 0, creating a floor effect.


```{r Verbal IQ ~ ADOS }

data_3 <- d3
data_3 <- subset(d3, d3$VerbalIQ > 0)

#SCALING
data_3$VerbalIQ_scale <- base::scale(data_3$VerbalIQ)
data_3$NonVerbalIQ_scale <- base::scale(data_3$NonVerbalIQ)
data_3$SocialIQ_scale <-base::scale(data_3$SocialIQ)
data_3$ADOS_scale <- base::scale(data_3$ADOS)

#DENSITY PLOT
rethinking::dens(data_3$VerbalIQ_scale)


#VERBAL IQ
M_IQ <- rethinking::map(
  alist(
    VerbalIQ_scale ~ dnorm (mu, sigma),
    mu <- a + b* ADOS_scale,
    a ~ dnorm (0, 5),
    b ~ dnorm (-0.1, 5), #expecting a negative relationship
    sigma ~ dunif (0, 5) 
  ), 
  data = data_3)

#MAP Values
precis(M_IQ)
#relatively small negative relationship it seems

#PLOTTING
#Plot 1 - Model with mean posterior values against actual data 
plot(VerbalIQ_scale ~ ADOS_scale, data = data_3)
abline(a=coef(M_IQ)["a"], b=coef(M_IQ)["b"])

#Plot 2
post <- extract.samples(M_IQ)
post[1:5,] #expecting five rows
#function
N <- 50
dN <- data_3[ 1:N , ]
# extract 20 samples from the posterior
post <- extract.samples( M_IQ , n=50 )
# display raw data and sample size
plot( dN$ADOS_scale , dN$VerbalIQ_scale,
    xlim=range(data_3$ADOS_scale) , ylim=range(data_3$VerbalIQ_scale) ,
    col=rangi2 , xlab="ADOS" , ylab="VerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )

#Okay, at this point we are reconsidering our linear choice, but we experience a floor effect, which might be possible to work around with a poisson model

models <- brms::bf("VerbalIQ_scale ~ ADOS")
brm_modelll <- brm(
  models, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = data_3, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_modelll, nsamples = 100)
#looking okay

```



``` {r Non-Verbal IQ ~ ADOS including TD}

N_IQ <- rethinking::map2stan(
  alist(
    NonVerbalIQ_scale ~ dnorm (mu, sigma),
    mu <- a + b* ADOS_scale,
    a ~ dnorm (0, 5),
    b ~ dnorm (-0.1, 5), #expecting a negative relationship
    sigma ~ dunif (0, 5) 
  ), 
  data = data_3)

#MAP Values
precis(N_IQ)
#our beta has both a great amount of values below and above zero, that doesn't look to good

#Plot 1 - Model with mean posterior values against actual data 
plot(NonVerbalIQ_scale ~ ADOS_scale, data = data_3)
abline(a=coef(N_IQ)["a"], b=coef(N_IQ)["b"])

#Plot 2
post <- extract.samples(N_IQ)
post[1:5,] #expecting five rows
#function
N <- 70
dN <- data_3[ 1:N , ]
# extract 20 samples from the posterior
post <- extract.samples( N_IQ , n=50 )
# display raw data and sample size
plot( dN$ADOS_scale , dN$NonVerbalIQ_scale,
    xlim=range(data_3$ADOS_scale) , ylim=range(data_3$NonVerbalIQ_scale) ,
    col=rangi2 , xlab="ADOS" , ylab="NonVerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )
#looks like it certainty about whether it is a positive or negative relationship is not given


n_model <- brms::bf("NonVerbalIQ_scale ~ ADOS")
brm_modelln <- brm(
  n_model, 
  prior = c(
    prior(normal(0, 5), class = "Intercept"),
    prior(normal(-0.1, 5), class = "b"), 
    prior(uniform(0, 5), class = "sigma")),
  data = data_3, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_modelln, nsamples = 100)
#looks good

```


``` {r Social IQ ~ ADOS including TD}
#SOCIAL IQ
S_IQ <- rethinking::map2stan(
  alist(
    SocialIQ_scale ~ dnorm (mu, sigma),
    mu <- a + b* ADOS_scale,
    a ~ dnorm (0, 5), #ADOS can't be below 0, and 
    b ~ dnorm (-0.1, 5), #expecting a negative relationship
    sigma ~ dunif (0, 5) 
  ), 
 data = data_3 )


#MAP Values
precis(S_IQ)
#rhat should approach from above 1 to 1 :) looking good

#Plot 1
plot(SocialIQ_scale ~ ADOS_scale, data = data_3)
abline(a=coef(S_IQ)["a"], b=coef(S_IQ)["b"])

#Plot 2
post <- extract.samples(S_IQ)
post[1:5,] #expecting five rows
#function
N <- 70
dN <- data_3[ 1:N , ]
# extract 20 samples from the posterior
post <- extract.samples( S_IQ , n=50 )
# display raw data and sample size
plot( dN$ADOS_scale , dN$SocialIQ_scale,
    xlim=range(data_3$ADOS_scale) , ylim=range(data_3$SocialIQ_scale) ,
    col=rangi2 , xlab="ADOS" , ylab="VerbalIQ" )
mtext(concat("N = ",N))
# plot the lines, with transparency
for ( i in 1:50 )
    abline( a=post$a[i] , b=post$b[i] , col=col.alpha("black",0.3) )

#once again brmsing for another pp_check
s_model <- brms::bf("SocialIQ ~ ADOS")
brm_modell <- brm(
  s_model, 
  prior = c(
    prior(normal(0, 100), class = "Intercept"),
    prior(normal(-0.1, 100), class = "b"), 
    prior(uniform(0, 100), class = "sigma")),
  data = data_3, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#Posterior Predictive Check
pp_check(brm_modell, nsamples = 100)
#looks okay :) 



```

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
```{r 4.1. Testing shared variance in IQ-scores}
#Testing correlation and covariates

#see 2.2 for correlations between measures in ASD kids

#TD kids
cor.test(d2$VerbalIQ, d2$SocialIQ, method = c("pearson", "kendall", "spearman"), exact = NULL, conf.level = 0.95)
0.20^2 #0.04
cor.test(d2$VerbalIQ, d2$NonVerbalIQ, method = c("pearson", "kendall", "spearman"), exact = NULL, conf.level = 0.95)
0.55^2 #0.3
cor.test(d2$NonVerbalIQ, d2$SocialIQ, method = c("pearson", "kendall", "spearman"), exact = NULL, conf.level = 0.95)
0.06^2 #0.004


```

4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.



5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


#NOTES TO SELF
if we were to compare models
simply use compare() and we'll get WAIC and stuff
plot(marginal_effects(m4, points = TRUE))
