---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Dumbledore's Army"
date: "December 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysing Heart Rate and Respiration data
```{r}
#Library
library(ggplot2)
library(pacman)
library(dplyr)
library(deSolve)
library(crqa)
library(caret)
library(tidyverse)
library(stringr)
library(stringi) 
library(pastecs) 
library(WRS2) 
library(sjPlot)
library(nlme) 
library(plyr) 
library(lmerTest)
p_load(tidyverse, stringr, Metrics, caret, lme4, simr, stats,lmerTest, stats, FinCal, PerformanceAnalytics, nonlinearTseries,purrr, grid.arrange)
p_load(tidyverse, crqa, readr, groupdata2, gridExtra, stringr, lmerTest, MuMIn)
p_load(deSolve)
p_load(crqa)
install.packages("tsbox")
library(tsbox)

library(tidyverse) ; library(stringr) ; library(stringi) ; library(pastecs) ; library(WRS2) ; library(sjPlot) ; library(nlme) ; library(plyr) ; library(lmerTest) ; library(pacman) 
p_load(tidyverse, stringr, Metrics, caret, lme4, simr, stats,lmerTest, stats, FinCal, PerformanceAnalytics, nonlinearTseries,purrr, grid.arrange)
p_load(tidyverse, crqa, readr, groupdata2, gridExtra, stringr, lmerTest, MuMIn)
p_load(deSolve)
p_load(crqa)

#Working directory
setwd("~/upgraded-lamp")

```


#Load the logfiles
```{r}

#picking out the 1st pair (G1)

t1 <- read.csv("CleanData/Study1_G1_T1_Synchronous.csv")
t2 <- read.csv("CleanData/Study1_G1_T2_TurnTaking.csv")
t3 <- read.csv("CleanData/Study1_G1_T3_SelfPaced.csv")
```


#Participant 1
Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
```{r setup, include=FALSE}

#Preparing the multiple plot function for later

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}


```


#CLEANING THE DATA in 3 steps
```{r}

#Step 1: Downsampling (slide 43)
p_load(groupdata2)

#downsampling T1
T1 = t1 %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))
#T2
T2 = t2 %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))
#T3
T3 = t3 %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

#Step 2: OUTLIER-REMOVAL
#defining function to remove outliers
removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm=T) +  
             (threshold*sd(ts,na.rm=T))) | 
       ts < (mean(ts,na.rm=T) - 
             (threshold*sd(ts,na.rm=T)))] = mean(ts,na.rm=T)  
  return(ts)}
threshold=2.5

#using function on the three trials
#T1
T1$HR1=removeOuts(T1$HR1,threshold)
T1$HR2=removeOuts(T1$HR2,threshold)
T1$Resp1=removeOuts(T1$Resp1,threshold)
T1$Resp2=removeOuts(T1$Resp2,threshold)
#T2
T2$HR1=removeOuts(T2$HR1,threshold)
T2$HR2=removeOuts(T2$HR2,threshold)
T2$Resp1=removeOuts(T2$Resp1,threshold)
T2$Resp2=removeOuts(T2$Resp2,threshold)
#T3
T3$HR1=removeOuts(T3$HR1,threshold)
T3$HR2=removeOuts(T3$HR2,threshold)
T3$Resp1=removeOuts(T3$Resp1,threshold)
T3$Resp2=removeOuts(T3$Resp2,threshold)


#Step 3: SCALING THE SIGNALS
#adding new column with scaled respiration
#T1
T1$Resp1S=scale(T1$Resp1) 
T1$Resp2S=scale(T1$Resp2)
T1$HR1S=scale(T1$HR1)
T1$HR2S=scale(T1$HR2)
#T2
T2$Resp1S=scale(T2$Resp1) 
T2$Resp2S=scale(T2$Resp2)
T2$HR1S=scale(T2$HR1)
T2$HR2S=scale(T2$HR2)
#T3
T3$Resp1S=scale(T3$Resp1) 
T3$Resp2S=scale(T3$Resp2)
T3$HR1S=scale(T3$HR1)
T3$HR2S=scale(T3$HR2)


#Now we can actually proceed
#HEARTRATE
# participant 1
HR1_T1 <- ggplot(T1, aes(x=time, y=HR1S))+geom_line()
HR1_T2 <- ggplot(T2, aes(x=time, y=HR1S))+geom_line()
HR1_T3 <- ggplot(T3, aes(x=time, y=HR1S))+geom_line()
#Displaying Participant 1, all Conditions, Heart rate
gridExtra::grid.arrange(HR1_T1, HR1_T2, HR1_T3)

#participant 2
HR2_T1 <- ggplot(T1, aes(x=time, y=HR2S))+geom_line()
HR2_T2 <- ggplot(T2, aes(x=time, y=HR2S))+geom_line()
HR2_T3 <- ggplot(T3, aes(x=time, y=HR2S))+geom_line()
#Displaying Participant 2, all Conditions, Heart rate
gridExtra::grid.arrange(HR2_T1, HR2_T2, HR2_T3)

#RESPIRATION
#participant 1
Resp1_T1 <- ggplot(T1, aes(x=time, y=Resp1S))+geom_line()
Resp1_T2 <- ggplot(T2, aes(x=time, y=Resp1S))+geom_line() #some of this data does not look like respiration...
   ggplot(t2, aes(x=time, y=Resp1))+geom_line() #comparing the raw version: flawed data even before
Resp1_T3 <- ggplot(T3, aes(x=time, y=Resp1S))+geom_line()
#Displaying Participant 1, all Conditions, Respiration data
gridExtra::grid.arrange(Resp1_T1, Resp1_T2, Resp1_T3)
#For Condition 1 and 2 we obviously have a problem, the data only has peaks, no lows as

#participant 2
Resp2_T1 <- ggplot(T1, aes(x=time, y=Resp2S))+geom_line()
Resp2_T2 <- ggplot(T2, aes(x=time, y=Resp2S))+geom_line()
Resp2_T3 <- ggplot(T3, aes(x=time, y=Resp2S))+geom_line()
#Particpant 2, all Conditions, Respiration
gridExtra::grid.arrange(Resp2_T1, Resp2_T2, Resp2_T3)

```
N.B: remember the slides: artifacts, downsampling, scaling.
N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)


#Eyeballing
- Can you eye-ball which condition if any displays more physiological coordination?
```{r}
#eyeballing material, heartrate
#three trials. QUESTION: Any difference in conditions between trials?
HR_T1 <- multiplot(HR1_T1, HR2_T1)
HR_T2 <- multiplot(HR1_T2, HR2_T2)
HR_T3 <- multiplot(HR1_T3, HR2_T3)


#eyeballing material, respiration
Resp_T1 <- multiplot(Resp1_T1, Resp2_T1)
Resp_T2 <- multiplot(Resp1_T2, Resp2_T2)
Resp_T3 <- multiplot(Resp1_T3, Resp2_T3)

#It seems like there is some coordination going on
#Let's keep investigating
```


#Crqa-ing! 
- Run crqa on heart rate and respiration data (find parameters, run crqa)
```{r}

#FINDING PARAMETERS (slide 77)
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE, fnnpercent = 10,  typeami = "mindip") #defining list

# Trial 1 - synchronous 
# Trial 2 - turn taking 
# Trial 3 - self paced 

# TRIAL 1
#Heart rate
ans_T1_HR = optimizeParam(T1$HR1S, T1$HR2S, par, min.rec = 3.5, max.rec = 4.5)
#radius=0.28, 
#emddim=20
#delay=0 (the smallest meaningful number is 1, so we'll use 1 in the following code) 

#Respiration
ans_T1_Resp = optimizeParam(T1$Resp1S, T1$Resp2S, par, min.rec = 3.5, max.rec = 4.5)
#radius=0.35
#emddim=2
#delay=19

#creating time series for crqa
T1_HR1 <- ts(T1$HR1S)
T1_HR2 <- ts(T1$HR2S)

results1_HR = crqa(T1_HR1, T1_HR2, delay=1, radius=0.28,embed=20, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
# RR = 3.56
# DET = 88.53
# TT = 6.29

#Times series for Respiration
T1_Resp1 <- ts(T1$Resp1S)
T1_Resp2 <- ts(T1$Resp2S)

results1_Resp = crqa(T1_Resp1, T1_Resp2, delay=19, radius=0.35,embed=2, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
# RR = 4.17
# DET = 83.86
# TT = 4.87


# TRIAL 2
#Heart rate
ans_T2_HR = optimizeParam(T2$HR1S, T2$HR2S, par, min.rec = 0, max.rec = 1)
#radius=1.47, 
#emddim=19
#delay=39 

#Respiration
ans_T2_Resp = optimizeParam(T2$Resp1S, T2$Resp2S, par, min.rec = 0, max.rec = 1)
#radius=0.67
#emddim=16
#delay=8

#creating time series for crqa
T2_HR1 <- ts(T2$HR1S)
T2_HR2 <- ts(T2$HR2S)

results2_HR = crqa(T2_HR1, T2_HR2, delay=39, embed=19, radius=1.47,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#it seems like crqa doesn't always work

T2_Resp1 <- ts(T2$Resp1S)
T2_Resp2 <- ts(T2$Resp2S)

results2_Resp = crqa(T2_Resp1, T2_Resp2, delay=8, radius=0.67,embed=16, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#it seems like crqa doesn't always work


# TRIAL 3
#Respiration
ans_T3_Resp = optimizeParam(T3$Resp1S, T3$Resp2S, par, min.rec = 3.5, max.rec = 4.5)
ans_T3_Resp
#radius=0.3502919
#emddim= 3
#delay= 47
T3_Resp1 <- ts(T3$Resp1S)
T3_Resp2 <- ts(T3$Resp2S)
#Results
results3_Resp = crqa(T3_Resp1, T3_Resp2, delay=47, radius=0.3502919,embed=3, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
results3_Resp
# RR = 0.4558156 
# DET = 62.67831
# TT = 3.689221

#Heartrate
ans_T3_HR = optimizeParam(T3$HR1S, T3$HR2S, par, min.rec = 1, max.rec = 5)
ans_T3_HR
#radius=2.731764
#emddim= 20
#delay= 12
#creating time series for crqa
T3_HR1 <- ts(T3$HR1S)
T3_HR2 <- ts(T3$HR2S)
#Results
results3_HR = crqa(T3_HR1, T3_HR2, delay=12, embed=20, radius=2.731764,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
results3_HR
# RR =   1.002546
# DET = 94.36765
# TT = 10.78701


#Instead of optimizing the parameters for every single trial, we want to find universal ones for Respiration and Heart rate for this Data Analysis and we can do so by taking the median of the different optimizedParameterd scores 


# Represent the plot 
#Trial 1
#Heart rate
RP1=results1_HR$RP
RP1 = matrix(as.numeric(RP1), nrow = ncol(RP1))
cols = c("white","blue4")
image(RP, xlab = "", ylab = "", col = cols)
#Respiration
RP2=results1_Resp$RP
RP2 = matrix(as.numeric(RP2), nrow = ncol(RP2))
cols = c("white","blue4")
image(RP2, xlab = "", ylab = "", col = cols)

#Trial 2
#Heart rate
RP3=results2_HR$RP
RP3 = matrix(as.numeric(RP3), nrow = ncol(RP3))
cols = c("white","blue4")
image(RP3, xlab = "", ylab = "", col = cols)
#Respiration
RP4=results2_Resp$RP
RP4 = matrix(as.numeric(RP4), nrow = ncol(RP4))
cols = c("white","blue4")
image(RP4, xlab = "", ylab = "", col = cols)

#Trial 3
#Heart rate
RP5=results3_HR$RP
RP5 = matrix(as.numeric(RP5), nrow = ncol(RP5))
cols = c("white","blue4")
image(RP5, xlab = "", ylab = "", col = cols)

#Respiration
RP6=results3_Resp$RP
RP6 = matrix(as.numeric(RP6), nrow = ncol(RP6))
cols = c("white","blue4")
image(RP6, xlab = "", ylab = "", col = cols)


# Explore the lags of coordination (plot of the time course of the diagonal lines)

# TRIAL 1 
#Respiration
Profile=drpdfromts(T1_Resp1, T1_Resp2,datatype = 'continuous',ws=50,radius=0.67)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial1_Resp <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')
# Heart rate 
Profile=drpdfromts(T1_HR1, T1_HR2,datatype = 'continuous',ws=50,radius=1.47)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial1_Heartrate <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')

multiplot(Trial1_Resp, Trial1_Heartrate)

#The delay peaks at 0 for both heart rate and respiration, which could be an indicator that they are very much synchronized (would make very much sense for the Synchronous condition!) and non is following the other but rather they are in sync.


# TRIAL 2
#Respiration
Profile=drpdfromts(T2_Resp1, T2_Resp2,datatype = 'continuous',ws=50,radius=0.67)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial2_Respiration <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')


# Heart rate 
Profile=drpdfromts(T2_HR1, T2_HR2,datatype = 'continuous',ws=50,radius=1.47)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial2_Heartrate <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')

multiplot(Trial2_Respiration, Trial2_Heartrate)

#It seems like one is leading and one is following the other since for both heart rate and respiration the delay is about 3 units away from 0 


# TRIAL 3 
#Respiration
Profile=drpdfromts(T3_Resp1, T3_Resp2,datatype = 'continuous',ws=50,radius=0.67)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial3_Respiration <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')


# Heart rate 
Profile=drpdfromts(T3_HR1, T3_HR2,datatype = 'continuous',ws=50,radius=1.47)
timecourse = round( seq(-5000,5000,100)/1000, digit = 1)
maxlag = Profile$maxlag/1000
profile = Profile$profile*100
Prof=data.frame(profile)
Trial3_Heartrate <- ggplot(Prof, aes(timecourse,profile))+geom_line()+ geom_vline(xintercept = timecourse[maxlag], colour='red')
Trial3_Heartrate

multiplot(Trial3_Respiration, Trial3_Heartrate)

#For respiration it peaks at 0, suggesting that most of the time they are very much in sync, whereas for heart rate one is following the other

```
Does this tell you more than just eyeballing the plots?
Yes, all in all, delay seems to be very different across Conditions, suggesting that the different conditions promote (as expected) different kinds of coordination: synchronization as well as one following the other

For TRIAL 1:
The delay peaks at 0 for both heart rate and respiration, which could be an indicator that they are very much synchronized (would make very much sense for the Synchronous condition!) and non is following the other but rather they are in sync.

For TRIAL 2:
It seems like one is leading and one is following the other since for both heart rate and respiration the delay is about 3 units away from 0 (one positive, one negative however)

For TRIAL 3:
For respiration it peaks at 0, suggesting that most of the time they are very much in sync, whereas for heart rate one is following the other


### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.
```{r}

#BIG FAT FUNCTION

#dependency for the preproz function
removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm=T) + (threshold*sd(ts,na.rm=T))) |
       ts < (mean(ts,na.rm=T) - (threshold*sd(ts,na.rm=T)))] =  mean(ts,na.rm=T)   
  return(ts)}
threshold=2.5 

setwd("/Users/lh/upgraded-lamp")
#function to downsample, rescale, remove outliers, print plots for HR and respiration as well as finding optimal parameters for crqa. Outputs a dataframe and prints plots
preproz = function(filename, graphs = T, noOutliers = T){ #requires the removeOuts function and a set threshold
  folder = "UnProcessed_OurData/"
  df = read_csv(paste(folder, filename, sep=""))
   
  #Downsample
  df = df %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(TimeMs= mean(TimeMs,na.rm=T), HR1 = mean(HR1,na.rm=T), HR2 = mean(HR2,na.rm=T), Resp1 =  mean(Resp1,na.rm=T), Resp2 = mean(Resp2,na.rm=T)) 
   
  if(noOutliers == T) {
  #Removing outliers
  df$HR1=removeOuts(df$HR1,threshold)
  df$HR2=removeOuts(df$HR2, threshold)
  df$Resp1=removeOuts(df$Resp1, threshold)
  df$Resp2=removeOuts(df$Resp2, threshold)
  }
  
  #Rescale
  df$Resp1S=scale(df$Resp1)  
  df$Resp2S=scale(df$Resp2)  
  df$HR1S=scale(df$HR1) 
  df$HR2S=scale(df$HR2)
  
  #Adding study identification colomns
  df$study = 3
  df$group = str_extract(filename, "G(\\d+)")
  df$trial = str_extract(filename, "T(\\d)")
  df$condition = gsub('.{4}$', '', strsplit(filename, "_")[[1]][4])
  df$filename = filename
  
  ###Calculating optimal parameters
  #setting parameters
  #We won't normaliza or rescale, since We've done that already - Mindiagline is set to 2, and seems fine Since it's a faily low starting point
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  
minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")

  #getting the optimal parameters and saving to opt_paramHR/Resp. Using try so it doesn't crash if there is an error
  
  #First getting parameters for HR data
  opt_paramHR = try(optimizeParam(df$HR1S, df$HR2S, par, min.rec = 2, max.rec = 8))
  #if the optimize function succeeded, save the results to the dataframe - if not, put NAs 
  if (length(opt_paramHR) > 2) {
    #need to unlist the parameters otherwise they can't be used for later calculations (no idea why it makes them a list in the first place)
   df$optRadiusHR = unlist(opt_paramHR[1])
   df$optEmbdimHR = unlist(opt_paramHR[2])
   df$optDelayHR = unlist(opt_paramHR[3])
   } else {
   df$optRadiusHR = NA
   df$optEmbdimHR = NA
   df$optDelayHR = NA
   }

  #Now for respiration data
  opt_paramResp = try(optimizeParam(df$Resp1S, df$Resp2S, par, min.rec = 2, max.rec = 8))
  #if the optimize function succeeded, save the results to the dataframe - if not, put NAs 
  if (length(opt_paramResp) > 2) {
    df$optRadiusResp = unlist(opt_paramResp[1])
    df$optEmbdimResp = unlist(opt_paramResp[2])
    df$optDelayResp = unlist(opt_paramResp[3])
    } else {
    df$optRadiusResp = NA
    df$optEmbdimResp = NA
    df$optDelayResp = NA
    }

  #creating new csv file with the new columns
  name = str_extract(filename, '.*(?=\\.csv)') #removing .csv from filename
  name = paste(name, "PROCESSED.csv", sep = '_')
  #creating new folder to store the files in 
  dir.create("preprocessed", showWarnings = FALSE) #stops warnings if folder already exists
  #writing file to the new folder
  write.csv(df, file.path("preprocessed", name), row.names=FALSE)

  if(graphs == T) {
  #Adding plots
  HR = ggplot(df, aes(TimeMs, HR1S)) + 
    geom_line() + 
    geom_line(aes(TimeMs, HR2S, color = "red")) + 
    ggtitle(filename) +
    theme(legend.position = "none")
  Resp = ggplot(df, aes(TimeMs, Resp1S)) + 
    geom_line() + 
    geom_line(aes(TimeMs, Resp2S, color = "red")) + 
    ggtitle(filename) +
    theme(legend.position = "none")

  #printing the arranged plots to console
  grid.arrange(HR, Resp)
  #grid.arrange can't be saved but will only print. Using arrangeGrob to save to variable and then to disk
  plots = arrangeGrob(HR, Resp) #generates g
  
  #creating a unique name for each plot based on the filename
  plotname = str_extract(filename, '.*(?=\\.csv)') #removing .csv from filename
  plotname = paste(plotname, "PLOT.png", sep = '_')
  #creating new folder to store the files in 
  dir.create("plots", showWarnings = FALSE) #stops warnings if folder already exists
  #writing file to the new folder
  #ggsave(plotname, plot = plots, path = "plots/")
  ggsave(file=plotname, plots, path = "plots/")
  }
  
  #Return df
  return(df)

}

files = list.files(path = "UnProcessed_OurData/", pattern = "*.csv")
processed = lapply(files, preproz)
processed = bind_rows(processed)

#Creating new column with optimal overall CRQA values
#Median over mean - More robust to outliers
processed = processed %>% 
  mutate(opt_dimHR = median(processed$optEmbdimHR, na.rm = T), 
         opt_delayHR = median(processed$optDelayHR, na.rm = T), 
         opt_radHR = median(processed$optRadiusHR, na.rm = T),
         opt_dimResp = median(processed$optEmbdimResp, na.rm = T), 
         opt_delayResp = median(processed$optDelayResp, na.rm = T), 
         opt_radResp = median(processed$optRadiusResp, na.rm = T))

write.csv(processed, file = "processedData.csv")

#setwd("/Users/FlowersnIce-cream/Google Drev/Hogwarts/R Studio/Alouishes/data")
#data <- read.csv("processedData.csv")

#If working with old data
#remove files 1,2, 4, 5 as well as all selfpaced conditions (Either Flatliners or a wierd condition used in the year before)
#If Working on the new data
# Remove G9_T2_Synchroneus which has flatlined Respiration

#removing files 1,2, 4, 5 as well as all selfpaced conditions

###REMOMVE 1, 2, 4, 5

```



#CRQA-ing on all Pre-processed data
- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.
```{r}


CRQA_final = data.frame()

PP_final = list.files('preprocessed', pattern = "*.csv", full.names = TRUE)

for (i in PP_final){ #PP_final
  #------ Read file ------
  file = read.csv(i, header = TRUE)
  
  #------ Extract info from filename ------
  Study = 3
  Group = str_extract(i, "G(\\d+)")
  Trial = str_extract(i, "T(\\d)")
  Condition = gsub('.{4}$', '', strsplit(i, "_")[[1]][4])
  
  #------ CRQA ------
  #Heart rate
  a_h = try(crqa(file$HR1S, file$HR2S,delay = 4 , embed= 20, radius= 2.328673,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  #If analysis fails, record NA
  if (length(a_h) > 2) {
    RR_h = a_h[1][[1]]
    DET_h = a_h[2][[1]] 
    NRLINE_h = a_h[3][[1]]
    maxL_h = a_h[4][[1]]
    L_h = a_h[5][[1]]
    ENTR_h = a_h[6][[1]]
    rENTR_h = a_h[7][[1]]
    LAM_h = a_h[8][[1]]
    TT_h = a_h[9][[1]]
  } else {
    RR_h = NA
    DET_h = NA
    NRLINE_h = NA
    maxL_h = NA
    L_h = NA
    ENTR_h = NA
    rENTR_h = NA
    LAM_h = NA
    TT_h = NA
  }
  #Respiration rate
  a_r = try(crqa(file$Resp1S, file$Resp2S,delay = 31, embed= 2, radius= 0.3845085,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  #If analysis fails, record NA
  if (length(a_r) > 2) {
    RR_r = a_r[1][[1]]
    DET_r = a_r[2][[1]]
    NRLINE_r = a_r[3][[1]]
    maxL_r = a_r[4][[1]]
    L_r = a_r[5][[1]]
    ENTR_r = a_r[6][[1]]
    rENTR_r = a_r[7][[1]]
    LAM_r = a_r[8][[1]]
    TT_r = a_r[9][[1]]
  } else {
    RR_r = NA
    DET_r = NA
    NRLINE_r = NA
    maxL_r = NA
    L_r = NA
    ENTR_r = NA
    rENTR_r = NA
    LAM_r = NA
    TT_r = NA
  }
  

  #------ Make dataframe ------  
  #Make relevant information into a one row data frame
  df_result = data.frame(Study, Group, Trial,
                             Condition, 
   RR_h, DET_h, NRLINE_h, maxL_h, L_h, ENTR_h, rENTR_h, LAM_h, TT_h, 
    RR_r, DET_r, NRLINE_r, maxL_r, L_r, ENTR_r, rENTR_r, LAM_r, TT_r
                             )
  
  #Bind the above row to the actual data set
  CRQA_final = rbind(CRQA_final, df_result)
  
}


#One option: We can compare the pairs with surrogate pairs. We are simply counting how many surrogate pairs have a higher synchronization than the pair, we$B!G(Bre investigating. All the surrogate pairs we consider $B!H(Bthe general population$B!I(B. This way, we are taking away the variation explained by having a functioning heart and doing a certain task

```


### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
```{r}

install.packages("tsbox")
library(tsbox)

#------ Shuffling ------
#function creating the shuffle files

pp_file_list = list.files('preprocessed', pattern = "*.csv", full.names = TRUE)
#Loop creating the shuffle files
for (i in pp_file_list){
  #read file
  file = read.csv(i, header = TRUE)
  
  #Shuffle variables
  HR1_x = sample(file$HR1S)
  file$HR1S = HR1_x
  
  HR2_x = sample(file$HR2S)
  file$HR2S = HR2_x
  
  Resp1_x = sample(file$Resp1S)
  file$Resp1S = Resp1_x
  
  Resp2_x = sample(file$Resp2S)
  file$Resp2S = Resp2_x
  
  #Write csv
  #Remove PP_data/ from name and add Shuffle_data
  name = str_extract(i, '.*(?=\\.csv)') #removing .csv from filename
  name = paste(name,"Shuffled.csv", sep = '_')
  write.csv(file, name, row.names = FALSE)
}

#### Put the New files in a new folder
#setwd("~/Downloads/")
#dir.create("shuffeled", showWarnings = FALSE) #stops warnings if folder already exists
#warnings()
#Doesn't work, move shuffeled files manually to shuffeled file


CRQA_final_shuffle = data.frame()

PP_final = list.files('shuffeled', pattern = "*.csv", full.names = TRUE)

for (i in PP_final){ #PP_final
  #------ Read file ------
  file = read.csv(i, header = TRUE)
  
  #------ Extract info from filename ------
  Study = 3
  Group = str_extract(i, "G(\\d+)")
  Trial = str_extract(i, "T(\\d)")
  Condition = gsub('.{4}$', '', strsplit(i, "_")[[1]][4])
  
  #------ CRQA ------
  #Heart rate
  a_h = try(crqa(file$HR1S, file$HR2S,delay = 4 , embed= 20, radius= 2.328673,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  #If analysis fails, record NA
  if (length(a_h) > 2) {
    RR_h = a_h[1][[1]]
    DET_h = a_h[2][[1]] 
    NRLINE_h = a_h[3][[1]]
    maxL_h = a_h[4][[1]]
    L_h = a_h[5][[1]]
    ENTR_h = a_h[6][[1]]
    rENTR_h = a_h[7][[1]]
    LAM_h = a_h[8][[1]]
    TT_h = a_h[9][[1]]
  } else {
    RR_h = NA
    DET_h = NA
    NRLINE_h = NA
    maxL_h = NA
    L_h = NA
    ENTR_h = NA
    rENTR_h = NA
    LAM_h = NA
    TT_h = NA
  }
  #Respiration rate
  a_r = try(crqa(file$Resp1S, file$Resp2S,delay = 31, embed= 2, radius= 0.3845085,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  #If analysis fails, record NA
  if (length(a_r) > 2) {
    RR_r = a_r[1][[1]]
    DET_r = a_r[2][[1]]
    NRLINE_r = a_r[3][[1]]
    maxL_r = a_r[4][[1]]
    L_r = a_r[5][[1]]
    ENTR_r = a_r[6][[1]]
    rENTR_r = a_r[7][[1]]
    LAM_r = a_r[8][[1]]
    TT_r = a_r[9][[1]]
  } else {
    RR_r = NA
    DET_r = NA
    NRLINE_r = NA
    maxL_r = NA
    L_r = NA
    ENTR_r = NA
    rENTR_r = NA
    LAM_r = NA
    TT_r = NA
  }
  

  #------ Make dataframe ------  
  #Make relevant information into a one row data frame
  df_result = data.frame(Study, Group, Trial,
                             Condition, 
   RR_h, DET_h, NRLINE_h, maxL_h, L_h, ENTR_h, rENTR_h, LAM_h, TT_h, 
    RR_r, DET_r, NRLINE_r, maxL_r, L_r, ENTR_r, rENTR_r, LAM_r, TT_r
                             )
  
  #Bind the above row to the actual data set
  CRQA_final_shuffle = rbind(CRQA_final_shuffle, df_result)
}

# In CRQA on our shuffeled data we still find some correlation between the pairs in the respiration - but we get very insignificant correlation in the heart rate data due to the muuuuuuuuch higher variability
#Some actually look down on Shuffeling, since U actually break/mess with the delayseries
``` 




##SURROGATE PAIRS
TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Malte will share a method to do this on screen.
 
One option: We can compare the pairs with surrogate pairs. We are simply counting how many surrogate pairs have a higher synchronization than the pair. This way, we are taking away the variation explained by having a functioning heart and doing a certain task
 
- Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
```{r}

data <- read.csv("wetransfer-88619a/processedData.csv")



One <- select(data, HR1S, Resp1S, condition, group)
One <- rename(One, c("HR1S"="HR", "Resp1S"="Resp"))
#Unique subject
One$subject <- ifelse(One$group=="G10", "10",
              ifelse(One$group == "G9", "9",
                     ifelse(One$group == "G8", "8",
                            ifelse(One$group == "G7", "7",
                                   ifelse(One$group == "G6", "6", 
                                          ifelse(One$group == "G5", "5",
                                                 ifelse(One$group == "G4", "4",
                                                        ifelse(One$group == "G3", "3",
                                                               ifelse(One$group == "G2", "2",
                                                                      ifelse(One$group == "G1", "1",
                                                                             NA))))))))))
#Malthe suggests case_when instead of Ifelse

Two <- select(data, HR2S, Resp2S, condition, group)
Two <- rename(Two, c("HR2S"="HR", "Resp2S"="Resp"))
#unique subject time 1
Two$subject <- ifelse(Two$group=="G10", "11",
              ifelse(Two$group == "G9", "12",
                     ifelse(Two$group == "G8", "13",
                            ifelse(Two$group == "G7", "14",
                                   ifelse(Two$group == "G6", "15", 
                                          ifelse(Two$group == "G5", "16",
                                                 ifelse(Two$group == "G4", "17",
                                                        ifelse(Two$group == "G3", "18",
                                                               ifelse(Two$group == "G2", "19",
                                                                      ifelse(Two$group == "G1", "20",
                                                                             NA))))))))))
AWESOME = rbind(One, Two)

#all combinations
AWESOME$subject <- as.numeric(AWESOME$subject)
combinations = expand.grid(p1 = unique(AWESOME$subject),p2 = unique(AWESOME$subject)) %>%
                             dplyr::filter(p1 < p2) # < makes sure that the same file doesn't go twice

surrogate_rqa = function(p1, p2, c) {
  d1 = filter(AWESOME, subject == p1, condition ==c)
  d2 = filter(AWESOME, subject == p2, condition ==c)
    #------ CRQA ------
  #Heart rate
  a_h = try(crqa(d1$HR, d2$HR,delay = 4 , embed= 20, radius= 2.328673,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  
  #If analysis fails, record NA
  if (length(a_h) > 2) {
    RR_h = a_h[1][[1]]
    DET_h = a_h[2][[1]] 
    NRLINE_h = a_h[3][[1]]
    maxL_h = a_h[4][[1]]
    L_h = a_h[5][[1]]
    ENTR_h = a_h[6][[1]]
    rENTR_h = a_h[7][[1]]
    LAM_h = a_h[8][[1]]
    TT_h = a_h[9][[1]]
  } else {
    RR_h = NA
    DET_h = NA
    NRLINE_h = NA
    maxL_h = NA
    L_h = NA
    ENTR_h = NA
    rENTR_h = NA
    LAM_h = NA
    TT_h = NA
  }
  #Respiration rate
  a_r = try(crqa(d1$Resp, d2$Resp,delay = 31, embed= 2, radius= 0.3845085,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))

  #If analysis fails, record NA
  if (length(a_r) > 2) {
    RR_r = a_r[1][[1]]
    DET_r = a_r[2][[1]]
    NRLINE_r = a_r[3][[1]]
    maxL_r = a_r[4][[1]]
    L_r = a_r[5][[1]]
    ENTR_r = a_r[6][[1]]
    rENTR_r = a_r[7][[1]]
    LAM_r = a_r[8][[1]]
    TT_r = a_r[9][[1]]
  } else {
    RR_r = NA
    DET_r = NA
    NRLINE_r = NA
    maxL_r = NA
    L_r = NA
    ENTR_r = NA
    rENTR_r = NA
    LAM_r = NA
    TT_r = NA
  }
df = data.frame(RR_h, DET_h, NRLINE_h, maxL_h, L_h, ENTR_h, rENTR_h, LAM_h, TT_h, RR_r, DET_r, NRLINE_r, maxL_r, L_r, ENTR_r, rENTR_r, LAM_r, TT_r)
return(df)
}

surrogate_rqa_sync = function(p1, p2) {
  surrogate_rqa(p1, p2, "Synchronous") %>%
    mutate(condition = "Synchronous")
}
surrogate_rqa_con = function(p1, p2) {
  surrogate_rqa(p1, p2, "Conversation")%>%
    mutate(condition = "Conversation")
}
surrogate_rqa_turn = function(p1, p2) {
  surrogate_rqa(p1, p2, "TurnTaking")%>%
    mutate(condition = "TurnTaking")
}

x <- combinations$p1
y <- combinations$p2

Sync <- map2_df(x, y, surrogate_rqa_sync)
Conv <- map2_df(x, y, surrogate_rqa_con)
Turn <- map2_df(x, y, surrogate_rqa_turn)
lol <- rbind(Sync, Conv, Turn)

write.csv(lol, "surrogates.csv", row.names = FALSE)


#Are we missing a t-test here
```
- Test whether crqa shows a difference between real and surrogate pairs
 

### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
```{r}

#we will look into the scores from the crqa

#BOXPLOTTING FOR RR, DET and MEAN LENGTH

# RR
ggplot(CRQA_final, aes(x=Condition,y = RR_r)) + geom_boxplot()
# check out how RR is much lower for conversation than the other two conditions - espeically synchronous. Yay!
  
ggplot(CRQA_final, aes(x=Condition,y = RR_h)) + geom_boxplot()
# Unexpected outcome: Relatively high scores in conversation-condition. Hm?
# variability for turntaking super wide for heart-rate???
# RR-scores of higher than heart-rate, which is the opposite of what we'd expect (lower variability for respiration relative to heartrate)

#Corresponding models
#Heart rate
crqa_final_model_h <- lmer(RR_h ~ Condition + (1|Group), CRQA_final)
summary(crqa_final_model_h)
#no t-values to check validity. We'll compare other model through anova
#unexpected: highest score for conversation

#Respiration
model_con <- lmer(RR_r ~ Condition +  (1|Group), CRQA_final)
summary(model_con)
# nice: RR is lower for conversation


  
#DETERMINISM
#Respiration
ggplot(CRQA_final, aes(x=Condition,y = DET_r)) + geom_boxplot()
# Conversation placed high again. Determinism-level similar across condition (amount of dots placed in diagonal lines)
#Heart rate
ggplot(CRQA_final, aes(x=Condition,y = DET_h)) + geom_boxplot()
#once more conversation tops synchronous

#Corresponding models
#Respiration
model_con_det_r <- lmer(DET_r ~ Condition +  (1|Group), CRQA_final)
summary(model_con_det_r)
#Heart rate
model_con_det_h <- lmer(DET_h~ Condition +  (1|Group), CRQA_final)
summary(model_con_det_h)



#MEAN LENGTH
#Respiration
ggplot(CRQA_final, aes(x=Condition,y = L_r)) + geom_boxplot()
#the synchronization remains for longer on average in the synchronous condition. Just like we'd expect. However, scores are quite low (3-4)
#Heart rate
ggplot(CRQA_final, aes(x=Condition,y = L_h)) + geom_boxplot()
# Not much variability between the conditions.
# NB: Mean length is surprisingly much higher for HR than respiration.
  
#Corresponding models
#Respiration
model_con_L_r <- lmer(L_r ~ Condition +  (1|Group), CRQA_final)
summary(model_con_L_r)
#Heart rate
model_con_L_h <- lmer(L_h ~ Condition +  (1|Group), CRQA_final)
summary(model_con_L_h)
#significant #woohoo


```



#Building interaction models
```{r}
setwd("~/upgraded-lamp")

#Reading surrogate data in
surro <- read.csv("surrogates.csv")

#creating an rbind with surrogates and real pairs
surro$Condition <- surro$condition
new_surro <- select(surro, Condition, RR_h, DET_h, L_h, RR_r, DET_r, L_r )
new_surro$pairs <- "surrogate"
new_real <- select(CRQA_final, Condition, RR_h, DET_h, L_h, RR_r, DET_r, L_r)
new_real$pairs <- "real"
new_real$Condition <- str_replace_all(new_real$Condition, c("TurnTa" = "TurnTaking", "Synchro" = "Synchronous", "Conversa" = "Conversation" ))
final_new_surro <- rbind(new_real, new_surro)


#BUILDING MODELS

#RR
#Respiration
model <- lm(RR_r ~ Condition * pairs, data = final_new_surro)
summary(model)

#Adjusted R-squared:  0.04039 and no significant p-values
#Heart rate
model_h <- lm(RR_h ~ Condition * pairs, data = final_new_surro)
summary(model_h)
#Adjusted R-squared:  0.04107 and p-values are far far far away from being significant


#DET
#Respiration
model_d <- lm(DET_r ~ Condition * pairs, data = final_new_surro)
summary(model_d)
#Adjusted R-squared:  0.03122 and no significant p-values
#Heart rate
model_d_h <- lm(DET_h ~ Condition * pairs, data = final_new_surro)
summary(model_d_h)
#Adjusted R-squared: 0.03977 and no significant p-values


#L
#Respiration
model_L_r <- lm(L_r ~ Condition * pairs, data = final_new_surro)
summary(model_L_r)
#Adjusted R-squared: 0.029 and no significant p-values
#Heart rate
model_L <- lm(L_h ~ Condition * pairs, data = final_new_surro)
summary(model_L)
#Adjusted R-squared: 0.05085  and no significant p-values


```
 
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
 
 In this experiment we looked for synchronization between people regarding 1) respiration and 2) heart rate. We have not, however, tested the assumption that these two measures are - at least somewhat - independent of each other. It could be, say, that respiration alone can predict heart rate. In this case, it would not make much sense for us to explore both measures.

We want to investigate if there is a link between coordination of respiration and coordination of heart rate. A simple way of testing this is comparing relevant crqa-scores for heart rate versus for respiration using a correlation-test. These scores are determinism [describe], recurrence rate [describe] and average length of diagonal lines (as a measure for how long the synchronization lasts).
This would result in three different results: One for each of the crqa-measures.

cor.test(CRQA_final$RR_HR, CRQA_final$RR_Resp) #recurrence rate
cor.test(CRQA_final$Det_HR, CRQA_final$Det_Resp) #determinism
cor.test(CRQA_final$L_HR, CRQA_final$L_Resp) #mean length

In order to have a baseline for comparison, the same tests could be run for the CRQA-scores coming from the surrogate pairs. If respiration coordination is a driver of heart rate coordination, we expect the correlation-scores to be more dramatic (further from 0) in the real pairs than in the surrogate-pairs.
