---
title: "Portfolio2Semester3"
author: "Dumbledore's Army: Astrid, Sebastian, Lisa, Fredrik, Lena"
date: "24 9 2018"
output: html_document
---

```{r}

#libraries

library(plyr)
library(ggplot2)
library(stringr)
library(lmerTest)


#Data

Data <- read.csv("First_Assignment_3_Semester.csv")
Data <- Data[c(2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18)] #getting rid of the X-column
Data <- rename(Data, c("nonVerbalIQ1"="nonVerbalIQ", "verbalIQ1"="verbalIQ"))


```


#Exercise 1
```{r, echo = FALSE}

#CREATING TWO GROUPS
Healthy <- subset(Data, Diagnosis=="TD")
Autists <- subset(Data, Diagnosis=="ASD")

#ADOS
#visualizing the two groups (hence x = Diagnosis) regarding the score on ADOS 
ggplot(Data, aes(x = Diagnosis, y = ADOS1)) + geom_boxplot()

t.test(Healthy$ADOS1 , Autists$ADOS1)

ggplot(Data, aes(x = ADOS1, fill = Diagnosis)) + geom_histogram()
#it seems there's a great variety within the autist group

#VERBALIQ and NONVERBALIQ
ls.str(Data)
#First we visualize the verbalIQ and nonVerbalIQ I'm in charge of
ggplot(Data, aes(x = Diagnosis, y = verbalIQ)) + geom_boxplot() 
ggplot(Data, aes(x = Diagnosis, y = nonVerbalIQ)) + geom_boxplot() 
#Verbal IQ seems asimilar - Let's compare  with a t-test
t.test(Healthy$verbalIQ,Autists$verbalIQ)
t.test(Healthy$nonVerbalIQ,Autists$nonVerbalIQ)
# p = 1.685e-05 vs. p = 0.1525
#As figured - verbalIQ seems to be a much stronger predictor for Autism than nonVerbalIQ
ggplot(Data, aes(x = Diagnosis, y = verbalIQ)) + geom_boxplot() + facet_wrap(~Gender)
ggplot(Data, aes(x = Diagnosis, y = nonVerbalIQ)) + geom_boxplot() + facet_wrap(~Gender)


#VISITS
t.test(Healthy$Visits,Autists$Visits)
#The two groups do not differ, suggesting that the visits between the two groups are relatively matched

ggplot(Data, aes(x = Diagnosis, y = Visits)) + geom_boxplot() 
#As for boxplot we see the obvious similarity between the two


#NUMBER OF WORDS USED BY MOTHERS
#testing number of words used (tokens) for mothers
t.test(Healthy$tokens_MOT,Autists$tokens_MOT)
#there is no significant difference between the outspokenness of the mothers of the two types of children. Let's see it in a boxplot
ggplot(Data, aes(x = Diagnosis, y = tokens_MOT)) + geom_boxplot() 
#again, we see a big overlap between the two categories. The reach for the ASD-condition is slightly wider, with less speech

#NUMBER OF WORDS USED BY CHILDREN
#testing number of words used (tokens) for children
t.test(Healthy$tokens_CHI,Autists$tokens_CHI)
#there is a significant difference in the amount of words used by normal versus autist children, with the normal children having a word count of +50% compared to the autists
ggplot(Data, aes(x = Diagnosis, y = tokens_CHI)) + geom_boxplot() 
#The box-plot is showing the distribution of autist children generally producing fewer words than normal children - although with a few outliers in the autists doing well


#NUMBER OF UNIQUE WORDS
types_mother <- t.test(Healthy$types_MOT,Autists$types_MOT)
types_mother
types_child <- t.test(Healthy$types_CHI,Autists$types_CHI)
types_child
#Boxplot
ggplot(Data, aes(x = Diagnosis, y = types_MOT)) + geom_boxplot() 
ggplot(Data, aes(x = Diagnosis, y = types_CHI)) + geom_boxplot()

#LENGTH OF UTTERANCES: CHI & MOT
MLU_mother <- t.test(Healthy$MOT_MLU,Autists$MOT_MLU)
MLU_mother
MLU_child <- t.test(Healthy$CHI_MLU,Autists$CHI_MLU)
MLU_child
#Boxplot
ggplot(Data, aes(x = Diagnosis, y = MOT_MLU)) + geom_boxplot()
ggplot(Data, aes(x = Diagnosis, y = CHI_MLU)) + geom_boxplot()


#GENDER, ETHNICITY, AGE
ggplot(Data, aes(x = Gender)) + geom_bar() + facet_wrap(~Diagnosis)
ggplot(Data, aes(x = Ethnicity,)) + geom_bar() + facet_wrap(~Diagnosis)
ggplot(Data, aes(x = Diagnosis, y = Age)) + geom_boxplot() 
#t.test
t.test(Healthy$Age,Autists$Age)
```

##Exercise 2: Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis
```{r, echo = FALSE}
library(lmerTest)

#BOXPLOT
boxplot(CHI_MLU ~ Diagnosis + Visits, col=c("brown","pink"),Data)

#CREATING A MODEL inlcuding Diagnosis and Visits (over time)
model_1 = lmer (CHI_MLU ~ Diagnosis + Visits + (1 + Visits| SUBJ), Data, REML = FALSE)
summary(model_1)

#NULL MODEL
null_model = lmer(CHI_MLU ~ Visits + (1 + Visits| SUBJ), Data, REML = FALSE)
summary (null_model)

#COMPARING NULL AND MODEL_1
anova(model_1, null_model)

#PLOTTING
ggplot(Data, aes(x=Visits, y=CHI_MLU, color=Diagnosis))+geom_point()+geom_smooth(method = "lm")

```
Diagnosis did not effect the MLU of the child as much as we thougth it would over time. Not different from the null_model which only includes Visits as a predictor
χ2(1)=2.28, p=0.13 (insignificant)

#Growth models
```{r, echo = FALSE}

#QUADRATIC
quadratic <- lmer(CHI_MLU ~ Visits + I(Visits^2) + Diagnosis + (1+Visits|SUBJ),Data, REML=FALSE)
summary(quadratic)
anova(null_model,quadratic)
ggplot(Data, aes(x=Visits, y=CHI_MLU, color=Diagnosis))+geom_point()+geom_smooth(method = "lm", formula = y ~ poly(x, 2))

#CUBIC
cubic <- lmer(CHI_MLU ~ Visits + I(Visits^2) + I(Visits^3) + Diagnosis + (1+Visits + I(Visits^2) + I(Visits^3)|SUBJ), Data, REML=FALSE)
summary(cubic)
anova(null_model,cubic)
ggplot(Data, aes(x=Visits, y=CHI_MLU, color=Diagnosis))+geom_point()+geom_smooth(method = "lm", formula = y ~ poly(x, 3))

#FINAL COMPARISON
anova(model_1, quadratic, cubic)

#BIC should be lower (Bayseian Information Criteria) -> How probable is it that we get this data with this model -> Is it really worth it adding these parameters/predictors
#We will use the awesome cubic model

```
Note to myself: growth models attempt to estimate between-person differences in within-person change. 


#Exercise 3
```{r }


#BOXPLOTTING
boxplot(MOT_MLU ~ Diagnosis + Visits, col=c("brown","pink"),Data)

#CREATING A MODEL
model_mot = lmer (MOT_MLU ~ Diagnosis + Visits + (1 + Visits| SUBJ), Data, REML = FALSE)
summary(model_mot)

#NULL MODEL
null_model_mot = lmer(MOT_MLU ~ Visits + (1 + Visits| SUBJ), Data, REML = FALSE)
summary (null_model_mot)

#COMPARING NULL AND REAL MODEL
anova(model_mot, null_model_mot)


#PLOTTING
ggplot(Data, aes(x=Visits, y=MOT_MLU, color=Diagnosis))+geom_point()+geom_smooth(method = "lm")

```



#Exercise 4
```{r, echo = FALSE}

#LOOKING FOR PREDICTORS
cor(Data$types_CHI, Data$tokens_CHI)
#very high correlation between these two. Supposedly we can just include one. We pick types. The word types are not as dependent on the mood of the kid on the day of visit

#FINAL MODEL
fresh_model = lmer(CHI_MLU ~ Diagnosis + Visits + types_CHI + verbalIQ + (1 + Visits|SUBJ), Data, REML=FALSE)
summary(fresh_model)


#FINAL TESTING
anova(fresh_model, null_model)

#very nice p-values and AIC-BIC
#note: is it OK that we compare our full model (with +4 predictors) to the null-model? Realistically, we cannot know, whether all the predictors have an influence (it could even be in the opposite direction)
#no malte, we no p-hack we got minds np

names(Data)
cor <- cor(Data[,9:15])
install.packages("corrplot")
library(corrplot)
corrplot(cor,method="shade", type="upper",tl.cex=0.7, title="Correlated Parameters")
round(cor, 2) #the number version

```
